Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-10-19 Paper.Rmd
rearranging plots and some stuffs..continued

* Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence difference (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations. Then we plotted mean of each measured quantity with standard error against Kendall's Tau of the noise copula family. Fig. \ref{fig_Cause4copula_C_kend} represents this picture for noise generating from a lower tail dependent Clayton copula.  
We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach (model selection method), here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise. (For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).

<!--
whereas for Fig. \ref{SM-fig_Cause4copula_C_spear} and Fig. \ref{SM-fig_Cause4copula_SC_spear} Spearman's Rho of noise copula were plotted along x-axis. -->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
\includegraphics[width=15 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/common_legend_cause4copula_stat.pdf}
\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e. Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison). p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae. \label{fig_Cause4copula_C_kend}}
\end{center}
\end{figure}

@@

\begin{equation}\label{eq.AS}
P_{i}(t+1) = b P_{i}(t) + g(\epsilon_{i}(t)) + h(\delta_{i}(t))
\end{equation}

$b$ was the linear growth rate. Global environmental fluctuation function $g(\epsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq. \ref{eq.AS_g}) where $\epsilon=(\epsilon_{1}, \epsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\epsilon}$.

\begin{align}
g(\epsilon) = \begin{cases} 
g_{1} = \begin{cases}
\epsilon & \text{if $\epsilon< 0$}; \\
0 & \text{otherwise}.
\end{cases}\\ \\
g_{2} = \begin{cases}
\epsilon & \text{if $\epsilon> 0$}; \\
0 & \text{otherwise}\label{eq.AS_g}.
\end{cases}
\end{cases} \text{where } \epsilon = N(\textbf{0},\Sigma_{\epsilon})\;\ \text{with} \;\ \Sigma_{\epsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
\end{align}

>>>>

* Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence difference (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations. Then we plotted mean of each measured quantity with standard error against Kendall's Tau of the noise copula family. Fig. \ref{fig_Cause4copula_C_kend} compares the statistics from population copula which was generated by noises drawn from lower tail dependent Clayton copula. It is possible to measure the statistics with variation of Spearman's Rho for any other noise copula family (e.g. see Fig. \ref{SM-fig_Cause4copula_SC_spear}). We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach (model selection method), here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise. (For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e. Spearman, Kendall, Pearson) and 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison). p-values $< 0.05$ (below horizontal purple line) indicates significantly similar estimate between noise and population copulae. \label{fig_Cause4copula_C_kend}}
\end{center}
\end{figure}

@@

\begin{equation}\label{eq.AS}
P_{i}(t+1) = b P_{i}(t) + g(\varepsilon_{i}(t)) + h(\delta_{i}(t))
\end{equation}

$b$ was the linear growth rate. Global environmental fluctuation function $g(\varepsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq. \ref{eq.AS_g}) where $\varepsilon=(\varepsilon_{1}, \varepsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\varepsilon}$.

\begin{align}
g(\varepsilon) = \begin{cases} 
g_{1} = \begin{cases}
\varepsilon & \text{if $\varepsilon< 0$}; \\
0 & \text{otherwise}.
\end{cases}\\ \\
g_{2} = \begin{cases}
\varepsilon & \text{if $\varepsilon> 0$}; \\
0 & \text{otherwise}\label{eq.AS_g}.
\end{cases}
\end{cases} \text{where } \varepsilon = N(\textbf{0},\Sigma_{\varepsilon})\;\ \text{with} \;\ \Sigma_{\varepsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
\end{align}

### 2018-10-19 SupportingInformation.Rmd
rearranging plots and some stuffs..continued

@@

plotter_asym_sens_b_varn<-function(mg,ml,a,b,numkeep_last,ploton){

  pop<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
  pop_d<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)

  neg_b<- -b
  pop_neg_b<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=neg_b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
  pop_d_neg_b<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=neg_b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)

  if(ploton==T){
    op<-par(mfrow=c(2,2),mar=c(4,3,1.5,1),mgp=c(2,1,0))

    #plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
    plot(pop[,1],pop[,2],col="red",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
    title(expression(paste("(A) g = ",g[1]," , ")), adj=0.4)
    mtext(paste("b = ",b),side = 3, line=0.2, adj=0.72, col="black")
    plot(pop_d[,1],pop_d[,2],col="blue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
    title(expression(paste("(B) g = ",g[2]," , ")), adj=0.4)
    mtext(paste("b = ",b),side = 3, line=0.2, adj=0.72, col="black")
    plot(pop_neg_b[,1],pop_neg_b[,2],col="orange",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
    title(expression(paste("(C) g = ",g[1]," , ")), adj=0.4)
    mtext(paste("b = ",neg_b),side = 3, line=0.2, adj=0.75, col="black")
    plot(pop_d_neg_b[,1],pop_d_neg_b[,2],col="skyblue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
    title(expression(paste("(D) g = ",g[2]," , ")), adj=0.4)
    mtext(paste("b = ",neg_b),side = 3, line=0.2, adj=0.75, col="black")

    par(op)
  }

  return(list(pop=pop,
              pop_d=pop_d,
              pop_neg_b=pop_neg_b,
              pop_d_neg_b=pop_d_neg_b))

@@

pdf("./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf")
ans_b_0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.1,numkeep_last=numkeep_last,ploton=T)
dev.off()

pdf("./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf")
ans_b_0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.9,numkeep_last=numkeep_last,ploton=T)
dev.off()

vlist_b_0.1_b_0.9<-c(ans_b_0.1,ans_b_0.9) # a list of 8 matrices

>>>>

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Spearman_vs_Spearman's Rho.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Kendall_vs_Spearman's Rho.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pearson_vs_Spearman's Rho.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Corl-Coru_vs_Spearman's Rho.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pl-Pu_vs_Spearman's Rho.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_D2u-D2l_vs_Spearman's Rho.pdf}
\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e. Spearman, Kendall, Pearson) and 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from upper-tail dependent \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison). p-values $< 0.05$ (below horizontal purple line) indicates significantly similar estimate between noise and population copulae. \label{fig_Cause4copula_SC_spear}}
\end{center}
\end{figure}

@@

plotter_asym_sens_b_varn<-function(mg,ml,a,b,numkeep_last,ploton,resloc){

  pop_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
  pop_d_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)

  if(ploton==T){
    pdf(paste0(resloc,"a_",a,"_b_",b,"_r_",r,"_asym_sens.pdf",sep=""),height=5,width=8)
    op<-par(mfrow=c(2,3),mar=c(4,4.2,1.5,1),mgp=c(2.5,0.6,0))
    #plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
    pop<-pop_g$pop
    geps<-pop_g$g_noise
    plot(mg[,1],geps[,1],col="red",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=1")
    plot(mg[,2],geps[,2],col="red",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=2")
    plot(pop[,1],pop[,2],col="red",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2)
    title(expression(paste("g = ",g[1]," , ")), adj=0.06,cex.main=1.5)
    mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9, col="black")

    pop_d<-pop_d_g$pop
    geps<-pop_d_g$g_noise
    plot(mg[,1],geps[,1],col="blue",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=1")
    plot(mg[,2],geps[,2],col="blue",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=2")
    plot(pop_d[,1],pop_d[,2],col="blue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2)
    title(expression(paste("g = ",g[2]," , ")), adj=0.06,cex.main=1.5)
    mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9, col="black")

    par(op)
    dev.off()
  }

  return(list(pop=pop,
              pop_d=pop_d))

@@

resloc<-"./Results/asym_sens_results/"
ans_b_0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
ans_b_neg0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
ans_b_0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.9,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
ans_b_neg0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.9,numkeep_last=numkeep_last,ploton=T,resloc=resloc)

vlist_b_0.1_b_0.9<-c(ans_b_0.1,ans_b_neg0.1,ans_b_0.9,ans_b_neg0.9) # a list of 8 matrices
